### 2023-12-12 main.tex
ss

We first look for the similarity between $\mathcal{H}_{x,y}$ and its transposed. Using \eqref{transpositionPropertyR} we obtain  
\begin{equation}\label{transpositionPropertyH}
    \mathcal{H}_{x,y}^{T}=\left(R_{x}R_{y}\right)\mathcal{H}_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the similarity transformation between $\mathcal{H}_{x,y}$ and its transposed. \\
We now look for a symmetry of $\mathcal{H}_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	\mathcal{H}_{x,y}S_{x,y}=S_{x,y}\mathcal{H}_{x,y}
	%\quad \Leftrightarrow \quad [\mathcal{H}_{x,y},S_{x,y}]=0
	\dt
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $\mathcal{H}_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C)$. Using the bilinearity of the coproduct operator and the fact that $C$ belongs to the centre of the Lie algebra, it is easy to show that for any matrix $E\in {gl}(N)$ the following holds:
\begin{equation}
 \text{if } \quad	[C,E]=0\qquad  \text{ then }\qquad \left[\Delta (C),\Delta(E)) \right]=0\dt
\end{equation}
We chose 
\begin{equation}
	E=\sum_{a=1}^{N-1}E_{aN}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta_{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta_{x,y}(E)$ denotes the coproduct acting on sites $x$ and $y$. As a consequence, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $\mathcal{H}_{x,y}$. Exploiting these considerations we may write

@@

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering {\color{blue}$\mathcal{A},\mathcal{B}\in {gl}(N)$}, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-\mathcal{B})}A\exp{(\mathcal{B})}=\mathcal{A}-\left[\mathcal{B},\mathcal{A}\right]+\frac{1}{2!}\left[B\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]-\frac{1}{3!}\left[\mathcal{B},\left[\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of this with $\mathcal{B}=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and with $\mathcal{A}$ equal to some of the generator $E_{CD}$ of the Lie algebra ${gl}(N)$ with $C,D\in\{1,\ldots,N\}$. Therefore we compute 
\begin{align}
	\left[\mathcal{B},\mathcal{A}\right]=\sum_{a=1}^{N-1}\left[E_{aN},E_{CD}\right]=\delta_{CN}\sum_{a=1}^{N-1}E_{aD}-(1-\delta_{DN})E_{CN}
\end{align}
and 
\begin{align}
	\left[\mathcal{B}\left[\mathcal{B},\mathcal{A}\right]\right]=&\left[\sum_{a=1}^{N-1}E_{aN},\delta_{CN}\sum_{b=1}^{N-1}E_{bD}-(1-\delta_{DN})E_{CN}\right]\nonumber\\=&
	\sum_{a=1}^{N-1}\left(\delta_{CN}\sum_{b=1}^{N-1}\left[E_{aN},E_{bD}\right]-(1-\delta_{DN})\left[E_{aN},E_{CN}\right]\right)\nonumber\\=&
	\sum_{a=1}^{N-1}\left(-\delta_{CN}\delta_{aD}\sum_{b=1}^{N-1}E_{bN}-(1-\delta_{DN})\delta_{CN}E_{aN}\right)\nonumber\\=&
	-\left(\delta_{CN}(1-\delta_{DN})+(1-\delta_{DN})\delta_{CN}\right)\sum_{b=1}^{N-1}E_{bN}\nonumber
	\\=&
	-2\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
\end{align}
From the third commutator on we always obtain zero, i.e. $[\mathcal{B},[\mathcal{B},[\mathcal{B},\mathcal{A}]]]=0$. All in all we have that 
 \begin{align}\label{HT-BA}
 	\exp{\left(-\sum_{a=1}^{N-1}E_{aN}\right)}E_{CD}\exp{\left(\sum_{a=1}^{N-1}E_{aN}\right)}=E_{CD}-\delta_{CN}\sum_{a=1}^{N-1}E_{aD}+(1-\delta_{DN})E_{CN}-\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
 \end{align}

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Section~\ref{subsection-description-process-LINE} with the hard-core constrain {\color{blue} (at most one particle per site)}, i.e. $\nu=1$. 
We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. 
We also recall the notation for the occupation variable which is called $\bm{\tau}=(\tau_1,\ldots,\tau_L)$ with $\tau_x\in\{1,\ldots,N\}$.
The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of ${gl}(N)$. The duality result stated in Section~\ref{statementDualitySubsection} is adapted to the present situation as follows. 
The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section~\ref{sectionIntegrabiliy}.\\

@@

$m$ species $a_{k}\in\{1,\ldots,N-1\}$  with $k=1,2,\ldots, m$. Then, the $m$-point correlations in the non-equilibrium steady state read

>>>>

We first look for the similarity between $\mathcal{H}_{x,y}$ and its transposed. Using \eqref{transpositionPropertyR} we obtain  
\begin{equation}\label{transpositionPropertyH}
    \mathcal{H}_{x,y}^{T}=\left(R_{x}R_{y}\right)\mathcal{H}_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the similarity transformation between $\mathcal{H}_{x,y}$ and its transposed. \\
We now look for a symmetry of $\mathcal{H}_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	\mathcal{H}_{x,y}S_{x,y}=S_{x,y}\mathcal{H}_{x,y}
	%\quad \Leftrightarrow \quad [\mathcal{H}_{x,y},S_{x,y}]=0
	\dt
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $\mathcal{H}_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C)$. Using the bilinearity of the coproduct operator and the fact that $C$ belongs to the centre of the Lie algebra, it is easy to show that for any linear combination of generators $E\in {gl}(N)$ the following holds:
\begin{equation}
 \text{if } \quad	[C,E]=0\qquad  \text{ then }\qquad \left[\Delta (C),\Delta(E)) \right]=0\dt
\end{equation}
Let
\begin{equation}
	E=\sum_{a=1}^{N-1}E_{aN}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta_{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta_{x,y}(E)$ denotes the coproduct acting on sites $x$ and $y$. As a consequence, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $\mathcal{H}_{x,y}$. Exploiting these considerations we may write

@@

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering {\color{blue}$\mathcal{A},\mathcal{B}\in {gl}(N)$}, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-\mathcal{B})}A\exp{(\mathcal{B})}=\mathcal{A}-\left[\mathcal{B},\mathcal{A}\right]+\frac{1}{2!}\left[\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]-\frac{1}{3!}\left[\mathcal{B},\left[\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of this with $\mathcal{B}=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and with $\mathcal{A}$ equal to some of the generator $E_{CD}$ of the Lie algebra ${gl}(N)$ with $C,D\in\{1,\ldots,N\}$. Therefore we compute 
\begin{align}
	\left[\mathcal{B},\mathcal{A}\right]=\sum_{a=1}^{N-1}\left[E_{aN},E_{CD}\right]=\delta_{CN}\sum_{a=1}^{N-1}E_{aD}-(1-\delta_{DN})E_{CN}
\end{align}
and 
\begin{align}
	\left[\mathcal{B}\left[\mathcal{B},\mathcal{A}\right]\right]=&\left[\sum_{a=1}^{N-1}E_{aN},\delta_{CN}\sum_{b=1}^{N-1}E_{bD}-(1-\delta_{DN})E_{CN}\right]\nonumber\\=&
	\sum_{a=1}^{N-1}\left(\delta_{CN}\sum_{b=1}^{N-1}\left[E_{aN},E_{bD}\right]-(1-\delta_{DN})\left[E_{aN},E_{CN}\right]\right)\nonumber\\=&
	\sum_{a=1}^{N-1}\left(-\delta_{CN}\delta_{aD}\sum_{b=1}^{N-1}E_{bN}-(1-\delta_{DN})\delta_{CN}E_{aN}\right)\nonumber\\=&
	-\left(\delta_{CN}(1-\delta_{DN})+(1-\delta_{DN})\delta_{CN}\right)\sum_{b=1}^{N-1}E_{bN}\nonumber
	\\=&
	-2\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
\end{align}
From the third commutator on we always obtain zero, i.e. $[\mathcal{B},[\mathcal{B},[\mathcal{B},\mathcal{A}]]]=0$. All in all we have that 
 \begin{align}\label{HT-BA}
 	\exp{\left(-\sum_{a=1}^{N-1}E_{aN}\right)}E_{CD}\exp{\left(\sum_{a=1}^{N-1}E_{aN}\right)}=E_{CD}-\delta_{CN}\sum_{a=1}^{N-1}E_{aD}+(1-\delta_{DN})E_{CN}-\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
 \end{align}

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Section~\ref{subsection-description-process-LINE} with the hard-core constrain {\color{blue} (at most one particle per site)}, i.e. $\nu=1$. 
We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. 
We also recall the notation for the occupation variable which is called $\bm{\tau}=(\tau_1,\ldots,\tau_L)$ with $\tau_x\in\{1,\ldots,N\}$.
The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of ${gl}(N)$. The duality result stated in Section~\ref{statementDualitySubsection} is adapted to the present situation as follows. 
The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{AB}^T= e_{BA}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section~\ref{sectionIntegrabiliy}.\\

@@

$m$ species $a_{k}\in\{1,\ldots,N-1\}$  with $k=1,2,\ldots, m$. Then, the $m$-point correlations in the non-equilibrium steady state read\todo{this I find still confusing}
